% abstract.tex — included by main.tex inside \begin{abstract}...\end{abstract}
Large language models are ranked and certified as safe on public benchmarks whose validity
rests on the benchmark not appearing in pre-training. We study \emph{benchmark contamination}
not as a measurement-hygiene problem but as a privacy/security vulnerability: contamination is a
visible symptom of memorization, and memorization is the mechanism by which sensitive content
leaks. Using the Pythia suite trained on the public Pile---so that membership is ground truth
rather than an inferred label---we run a systematic, pre-registered comparison of existing
contamination/membership detectors (LOSS, Min-K\%, Min-K\%++, zlib) against a per-item
\emph{extraction} outcome. We make no claim to a new detector or metric. Our contribution is a
\emph{controlled} result: a pre-registered partial-correlation and mediation analysis that
isolates the role of raw per-item loss. We find that the apparent contamination$\rightarrow$leakage
association is \emph{loss-mediated to the resolution of this experiment}: the calibrated
reference-free detectors---which are themselves strongly-to-moderately collinear with loss
(Spearman $0.74$--$0.90$)---add no positive predictive value beyond it, and their residual partials
are null or, for the most loss-collinear detector (Min-K\%), weakly negative in a manner consistent
with a suppression artifact rather than substantive inverse prediction. This absence of positive
residual survives a non-linear (cubic-residual and decile-stratified) loss control and
deduplication, and is not explained by token frequency or the zero-inflated outcome. We frame this
as a \emph{membership-detection-versus-leakage-prediction divergence}: the detectors the field
optimizes for membership are not the right instrument for the privacy question. \textbf{These results are preliminary,
obtained on the smallest ($160$M) Pythia model on CPU; the pipeline is built so that the
GPU-scale replication is a single configuration change.} All analyses are pre-registered and every
number is reproducible from a seeded script.
