% =============================================================================
% evaluation.tex  --  reconciled "Evaluation Overview".
% Owner: Subagent A, reconciled with docs/experiment_design.md + method memo.
%
% MAJOR DIVERGENCES FROM THE PASTED DRAFT (each flagged; please review):
%  [D1] REMOVED the "Proposed Method: Internal Activation Analysis" as a claimed
%       contribution. The advisor constraint forbids proposing a novel detector. It
%       survives only as an optional exploratory probe in the Discussion, not a result.
%  [D2] REPLACED Precision/Recall/F1 as primary metrics with TPR@low-FPR + log-scale
%       ROC + AUC (Carlini S&P'22 convention). P/R/F1 retained only as a secondary,
%       fixed-threshold view for the benchmark-flagging use case.
%  [D3] REMOVED the "fine-tune to force overfitting on PII" positive control. We do not
%       train models. Memorization ground truth comes from public Pile membership and
%       Pythia's known duplication counts / checkpoints. Enron PII is studied because
%       the Enron Emails corpus is itself a Pile subset (i.e., already in training),
%       giving controlled PII ground truth without inducing new memorization.
%       [VERIFY] confirm Enron Emails is a Pile component before citing as such.
%  [D4] REPLACED k-fold cross-validation (a trained-classifier validation scheme) with
%       multi-seed bootstrap confidence intervals + a permutation test, appropriate for
%       threshold-free attack metrics.
% =============================================================================

\section{Evaluation Overview}
\label{sec:eval}

\subsection{Threat model and success criteria}
We frame contamination detection as a membership/exposure attack with an explicit
adversary (Section omitted here; see \texttt{docs/experiment\_design.md}). Goals range
from membership inference on a single item, to benchmark-level contamination
confirmation, to verbatim extraction and PII leakage. Each detector is evaluated at its
minimum access tier (gray-box logprobs for LOSS/Min-K\%/zlib; white-box logits for
Min-K\%++). Success is defined by the security-appropriate operating point rather than
average accuracy.

\subsection{Methods under comparison}
We evaluate \emph{existing} detectors only---we propose no new detector. The
per-item membership suite is LOSS/perplexity~\cite{yeom2018privacy}, Min-K\%
Prob~\cite{shi2024detecting}, Min-K\%++~\cite{zhang2025minkpp}, and the zlib-entropy
ratio~\cite{carlini2021extracting}. Two further tests operate off the per-item
likelihood axis: corpus-side $n$-gram overlap~\cite{brown2020gpt3}, a model-free
data-side check used to construct ground-truth contamination labels for benchmark items,
and the Oren permutation/exchangeability test~\cite{oren2024proving}, a benchmark-level
test that compares the canonical ordering of a benchmark's examples against random
shufflings to certify contamination with a controlled false-positive rate. The leakage
outcome is prefix-continuation extractable memorization under greedy
decoding~\cite{carlini2023quantifying}; on the controlled corpus we additionally measure
regex-detected PII leakage, framed via the PII-leakage games of
Lukas~et~al.~\cite{lukas2023pii}. Related approaches we deliberately \emph{do not}
evaluate---guided prompting~\cite{golchin2024timetravel}, neighbourhood and shadow-model
reference attacks~\cite{mattern2023neighbourhood,shokri2017membership}, and the
divergence-style extraction of production
models~\cite{nasr2025scalable}---are discussed in Section~\ref{sec:relatedwork}.
\textbf{[D1]} An internal-activation probe is reported, if at all, only as exploratory
analysis in the Discussion, not as a contribution.

\subsection{Data}
Table~\ref{tab:datasets} summarizes every corpus and benchmark used or referenced below.

\paragraph{Models and corpus.} The primary model is the Pythia
suite~\cite{biderman2023pythia}, trained on the public Pile~\cite{gao2020pile}; its
reconstructible training order, $154$ checkpoints, multiple sizes, and deduplicated
variant provide exact membership ground truth. We use the released MIMIR member/
non-member splits~\cite{duan2024mia}, which control $n$-gram overlap between members and
non-members. OLMo~\cite{groeneveld2024olmo} on Dolma~\cite{soldaini2024dolma} is a
secondary replication target. The Pile sits within the broader weakly filtered
web-scrape regime---Common Crawl~\cite{commoncrawl} and its filtered derivatives
C4~\cite{raffel2020c4,dodge2021c4} and RedPajama~\cite{weber2024redpajama}---that makes
benchmark contamination structural rather than adversarial.

\input{datasets_table}

\paragraph{Benchmarks and PII.} Contamination is tested against MMLU, GSM8K, HumanEval,
HellaSwag, TruthfulQA, and BoolQ. \textbf{[D3]} For PII leakage we use the Enron Emails
data \emph{as a Pile subset already present in Pythia's training data}, plus a synthetic
PII set for controlled structure, rather than fine-tuning a model to memorize PII. All
PII results are reported in aggregate; no real PII is reproduced in the paper.

\subsection{Metrics (each justified)}
\textbf{[D2]} Following the membership-inference-from-first-principles
convention~\cite{carlini2022lira}, the primary metric is \emph{true-positive rate at a
fixed low false-positive rate} (TPR @ $0.1\%$ and $1\%$ FPR) reported with
\emph{log-scale ROC}; AUC-ROC is reported secondarily. These capture whether a detector
\emph{confidently} identifies members---the privacy-relevant regime---which average-case
accuracy hides. For benchmark flagging at a chosen operating threshold we additionally
report precision/recall/F1 as a secondary, application-facing view. The leakage outcome
is the \emph{extraction rate}~\cite{carlini2023quantifying}. The headline analysis is
the \emph{Spearman correlation between per-item contamination score and per-item
extraction/leakage outcome}, with bootstrap confidence intervals and a pre-registered
partial-correlation control that isolates the contribution of raw loss---the quantitative
form of the paper's central question.

\subsection{Validation and controls}
\textbf{[D4]} Robustness is established by repeating each measurement over multiple
seeds with bootstrap confidence intervals on TPR@FPR and on the Spearman correlation,
and by a permutation/exchangeability test for benchmark-level
contamination~\cite{oren2024proving}. We include ablations that preempt the standard
confounds: deduplicated versus non-deduplicated Pythia (duplication), frequency-matched
member/non-member splits (string frequency), and model-size scaling (does the
contamination$\rightarrow$leakage link strengthen with scale, as memorization
does~\cite{carlini2023quantifying}). Differentially private
training~\cite{abadi2016deep,li2022dpllm} is discussed as the mitigation direction
(Section~\ref{sec:dp}), not implemented, since it is a producer-side defense applied at
training time rather than an auditor-side detector.

This section fixes the threat model, methods, data, and metrics; the empirical results
under this protocol---per-detector TPR at low FPR with log-scale ROC, extraction rates,
and the headline contamination$\rightarrow$leakage correlation with confidence
intervals---are reported in the results section, with every reported number tracing to a
logged harness run.
