% =============================================================================
% introduction.tex -- Introduction + explicit honest contributions.
% Owner: Subagent L (Deep Literature Research & Lit-Review Finalizer).
% All \cite keys -> ../references.bib.
%
% SCOPE CONTRACT (mirrors docs/method_selection_memo.md S5): the contribution is a
% security REFRAMING + comparative evaluation of EXISTING detectors, NOT a novel
% detector or metric. Every method named as evaluated is in the implemented set
% (LOSS, Min-K%, Min-K%++, zlib, n-gram overlap, Oren permutation test, extractable
% memorization, regex PII on Enron-in-Pile). Everything else is "related, not
% evaluated."
% =============================================================================

\section{Introduction}
\label{sec:intro}

Large language models (LLMs) are ranked, selected, and certified as safe largely on
the basis of their scores on public benchmarks~\cite{hendrycks2021mmlu,cobbe2021gsm8k}.
Those scores are only meaningful under one assumption: that the evaluation data was
absent from pre-training. The assumption is increasingly untenable. Benchmarks are
small, static, and endlessly redistributed across the web, while training corpora are
weakly filtered crawls assembled at the scale of hundreds of gigabytes to
petabytes~\cite{commoncrawl,gao2020pile}; benchmark items are therefore swept into the
next crawl by ordinary copying, with no adversary required. The resulting
\emph{benchmark contamination}, the presence of evaluation data in the training
corpus~\cite{golchin2024timetravel}, is usually treated as a measurement-hygiene
problem: a contaminated score over-states capability~\cite{ravaut2024survey}.

We argue that contamination is better understood as a \emph{privacy and security}
vulnerability, and we study it as one. The same over-parameterized models that score
highly on a leaked benchmark also memorize and can regurgitate verbatim training
sequences, including personally identifiable information (PII) that co-occurs in the
same corpora~\cite{carlini2021extracting,carlini2023quantifying}. Contamination, in
this view, is a visible symptom of unintended memorization, and memorization is the
mechanism by which sensitive content leaks. If a cheap, model-side contamination signal
predicts which items the model has memorized, then the act of contaminating a benchmark
is not merely inflating a metric; it is exposing a leakage channel. We make this
contamination~$\rightarrow$~memorization~$\rightarrow$~leakage chain the object of
empirical study, on models whose training corpus is fully public so that membership is
ground truth rather than a guess.

The privacy/security community has, however, established that membership signal on
pre-trained LLMs is weak: large-scale audits on the Pythia suite and The Pile report
that membership-inference attacks (MIAs) barely exceed chance, and that apparent
successes often reflect distribution shift between the member and non-member sets rather
than membership itself~\cite{duan2024mia}. We take this finding as a constraint, not an
obstacle. Rather than claim a stronger attack, we ask a sharper, security-relevant
question: \emph{even where the membership signal is weak, does it still predict concrete
leakage?} Answering it requires the evaluation discipline that security venues expect of
a privacy attack, true-positive rate at a low, fixed false-positive rate, read off a
log-scale ROC curve, rather than an average-case AUC that hides whether the attack ever
fires confidently~\cite{carlini2022lira}. It also exposes a question the
membership-inference literature does not ask: detectors are tuned and ranked by how well
they separate members from non-members, but leakage is a property of \emph{how much} the
model memorized a specific item. We therefore evaluate each detector not only as a
membership classifier but as a predictor of concrete leakage, and ask whether the two
objectives coincide, finding that they do not.

\paragraph{Contributions (and explicit non-contributions).}
We are deliberate about what this paper is and is not. It is \emph{not} a new detector,
attack, or metric: every detection method we run is from prior
work~\cite{yeom2018privacy,shi2024detecting,zhang2025minkpp,carlini2021extracting,brown2020gpt3,oren2024proving},
and our evaluation protocol is the established low-FPR convention of
Carlini~et~al.~\cite{carlini2022lira}. Within that honest scope, our contributions are:
\begin{itemize}
 \item \textbf{A security reframing and threat model.} We recast benchmark
 contamination as a membership/exposure vulnerability with an explicit adversary and
 graded goals, membership inference on a single item, benchmark-level contamination
 confirmation, and verbatim/PII extraction, rather than as a measurement artifact
 (Section~\ref{sec:background}, Section~\ref{sec:eval}).
 \item \textbf{A systematic comparative evaluation of existing detectors under the S\&P
 low-FPR protocol on ground-truth Pile membership.} We evaluate LOSS/perplexity,
 Min-K\%, Min-K\%++, and the zlib ratio as membership detectors, the corpus-side
 $n$-gram overlap test as a contamination-label oracle, and the Oren
 permutation/exchangeability test at the benchmark level, all on Pythia trained on the
 public Pile, reporting TPR at $0.1\%$ and $1\%$ FPR with log-scale ROC and bootstrap
 confidence intervals, with explicit controls for the frequency, duplication, and
 temporal confounds that prior work
 identifies~\cite{biderman2023pythia,gao2020pile,duan2024mia}.
 \item \textbf{A pre-registered measurement of \emph{which} contamination signal predicts
 leakage, and which does not.} We correlate per-item contamination scores against an
 extraction outcome, prefix-continuation extractable memorization under greedy
 decoding~\cite{carlini2023quantifying}, and, on the Enron Emails subset that already
 sits inside the Pile, against regex-detected PII leakage~\cite{lukas2023pii}. A
 pre-registered partial-correlation and mediation control then isolates the role of raw
 loss. In our ground-truth $160$M-parameter setting we find that the
 contamination$\rightarrow$leakage association is \emph{loss-mediated to the resolution of
 this experiment}: once loss is held fixed, the calibrated reference-free detectors
 (Min-K\%, Min-K\%++, zlib) add no positive predictive value. These detectors are
 themselves near-collinear transforms of loss (Spearman $0.74$--$0.90$), so we read a
 negative residual for the most collinear of them (Min-K\%) as a likely suppression
 artifact rather than substantive inverse prediction, and claim only the conservative
 result: no positive signal beyond loss. The calibrations that improve membership-detection
 AUC thus do not retain the loss-magnitude signal that predicts leakage, a divergence
 between membership detection and leakage prediction that we report as our central empirical
 finding (robust to deduplication and a non-linear loss control, and not explained by token
 frequency or the zero-inflated outcome).
\end{itemize}
We do not propose internal-probe or other novel detectors as contributions, do not train
or fine-tune models, and do not attack closed production systems for real third-party
PII; differential privacy and related defenses are discussed as the mitigation direction
only.
