% threat_model.tex — standalone Threat Model section (promoted from a subsection).
\section{Threat Model}
\label{sec:threat}

We frame contamination detection as a membership/exposure attack and state the adversary
explicitly, following the convention that a privacy attack must be evaluated by its behaviour at a
low false-positive operating point rather than on average~\cite{carlini2022lira}.

\paragraph{Adversary goals (graded).}
\begin{itemize}
  \item \textbf{G1 --- Membership inference.} Decide whether a specific sequence (a benchmark item,
  document, or record) was in the training corpus.
  \item \textbf{G2 --- Benchmark-level contamination confirmation.} Decide, with a controlled
  false-positive rate, whether an entire benchmark was trained on.
  \item \textbf{G3 --- Extraction / leakage.} Recover verbatim content---and, on a controlled
  corpus, PII---that was in training. This is the concrete harm; G1--G2 are of interest largely
  insofar as they predict G3.
\end{itemize}

\paragraph{Adversary knowledge and access.} We grade detectors by the minimum access each requires:
\emph{black-box} (text in, text out; e.g.\ guided prompting, which we do not evaluate),
\emph{gray-box} (per-token log-probabilities / loss; LOSS, Min-K\%, zlib), and \emph{white-box}
(the full next-token distribution; Min-K\%++). The corpus-side $n$-gram test instead assumes access
to the training corpus (available here because the Pile is public) and is used to construct
ground-truth contamination labels, not as a model-access attack. For our ground-truth experiments
the \emph{auditor} additionally knows the public training corpus; the modelled attacker does not
need corpus access for G1/G3.

\paragraph{Success criteria.} G1: true-positive rate at $0.1\%$ and $1\%$ false-positive rate
(log-scale ROC), with AUC secondary and bootstrap confidence intervals. G2: a permutation-test
$p$-value below threshold with a controlled false-positive rate~\cite{oren2024proving}. G3: a
non-zero extraction rate, and---our headline analysis---a positive association between a per-item
contamination score and the per-item extraction outcome that \emph{survives controlling for raw
loss}. The last criterion is what distinguishes a contamination signal that genuinely predicts
leakage from one that merely restates the model's loss.

\paragraph{Out of scope.} We do not attack closed production models for real third-party PII, do not
train or fine-tune models, and propose no new detector. Differential privacy is discussed as the
producer-side mitigation our threat model motivates (Section~\ref{sec:dp}), not implemented.
